%Thedore Ian Martiny
%Summer 2017 Homework 6

\documentclass[addpoints]{exam}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\reals}{\mathbb{R}}
\renewcommand{\baselinestretch}{1.5}
%\setlength{\textwidth}{16cm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wasysym}
\usepackage{xcolor}
 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\definecolor{Gray}{gray}{0.85}
 
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\newcolumntype{a}{>{\columncolor{Gray}}c} 
\lstset{style=mystyle}
\begin{document}
\singlespacing

\begin{center}
  {\large\textbf{CSCI 2824 - Discrete Structures}}

  {\large\textbf{Homework 6}}
\end{center}

You MUST show your work. If you only present answers you will receive minimal
credit. This homework is worth \numpoints pts.

\textbf{Due: Wednesday July 19}

\begin{questions}
  \question For the following question consider rolling two dice. One dice is
  red and one is blue, but they are otherwise normal dice.
  \begin{parts}
    \part[4] How many outcomes result in a sum of 2? In a sum of 12?
    \begin{solution}
      Only one outcome results in a sum of 2, 1 and 1.

      Only one outcome results in a sum of 12, 6 and 6.
    \end{solution}

    \part[4] How many outcomes have the blue die showing 2?
    \begin{solution}
      Six outcomes have the blue die showing 2. Any roll with the red die (of which there are 6).
    \end{solution}

    \part[4] How many outcomes give an even sum?
    \begin{solution}
      Possible even sums are 2, 4, 6, 8, 10, 12. 

      There is only one way to make 2.

      There are 3 ways to make 4, 2 and 2, red 3 blue 1 and red 1 blue 3.

      There are 5 ways to make 6, 3 and 3, red 4 blue 2, red 2 blue 4, red 5
      blue 1, and red 1 blue 5.

      There are 5 ways to make 8, 4 and 4, red 5 blue 3, red 3 blue 5, red 6
      blue 2, and red 2 blue 6.

      There are 3 ways to make 10, 5 and 5, red 6 blue 4, and red 4 blue 6.

      There is only one way to make 12.

      Altogether there are 18 ways to get an even sum.
    \end{solution}
  \end{parts}

  \question For the following questions consider an eight-bit string.
  \begin{parts}
    \part[2] How many eight-bit strings begin with $1100$?
    \begin{solution}
      We know the first four bits of the eight-bit string, the last 4 have 2
      choices each. There are $2^4 =16$ eight-bit strings with $1100$ in the
      first four bits.
    \end{solution}

    \part[2] How many eight-bit strings have exactly one $1$ in them?
    \begin{solution}
      If there is exactly one $1$ then the other $7$ bits are all $0$. Meaning
      we only have to choose which place to put the $1$ in. There are
      $\binom{8}{1} = 8$ ways to place the $1$.
    \end{solution}

    \part[5] How many eight-bit strings have exactly two $1$'s in them?
    \begin{solution}
      We know there will be two $1$s and six $0$s the only question is where to
      place the two $1$s. There are $\binom{8}{2} = 28$ ways to place the two
      $1$s.
    \end{solution}

    \part[5] How many eight-bit strings have at least one $1$ in them?
    \begin{solution}
      The long way to do this problem is to compute how many strings have one
      $1$, how many have two $1s$ etc. The shorter is to compute the opposite.
      How many have no $1$s. There is only one eight-bit string which has no
      $1$s, every other string has at least one.

      Thus there are $2^8-1 = 255$ eight-bit string which have at least one $1$
      in them.
    \end{solution}
  \end{parts}

  \question For the following questions determine how many permutations
  (orderings) can be formed from $\{A,B,C,\allowbreak D,E\}$ subject to the
  given constraints.
  \begin{parts}
    \part[4] The ordering contains the substring $ACE$ (in that order directly,
    as in $ACEDB$ is one, but $ADCBE$ is not)
    \begin{solution}
      First we compute how many places can $ACE$ go. It could start at the
      first, second, or third character. Meaning there are three ways of
      choosing where to place $ACE$.

      There are 2 ways of ordering the remaining two letters. 

      Thus altogether we have $3\cdot 2= 6$ ways of forming strings with $ACE$
      as a substring.
    \end{solution}

    \part[6] Does not contain the substring $AB$ nor $CD$ (again directly,
    $ACEDB$ is acceptable).
    \begin{solution}
      To compute this we work with the opposite. That is we compute how many
      strings have $AB$, have $CD$ and have $AB$ AND $CD$ and then subtract that
      from the total number of strings. Note we have to compute the strings that
      have both $AB$ and $CD$ and not do it separately so that we don't double
      count some strings.

      To count the number of strings that contain just $AB$ is just $4!$ since
      we have to permute 4 objects: $AB, C, D, E$.

      Similarly to count the number of strings that contain just $CD$ is $4!$ as
      well.

      To count the strings that contain both $AB$ and $CD$ is $3!$ since we
      are permuting 3 object, $AB, CD, E$.

      There are $5! = 120$ total strings from $\{A,B,C,D,E\}$ so there are $120
      - (24 + 24 - 6) = 78$ strings which do not contain $AB$ nor $CD$ as
      substrings. We need to subtract the 6, is because adding the two separate
      cases double counts the cases where they both show up.
    \end{solution}

    \part[10] $A$ appears before $C$ which appears before $E$ (here there can be
    letters between).
    \begin{solution}
      For this problem we will recognize why part of the total strings of length
      5 are acceptable. There are $5! = 120$ possible strings. The possible
      cases (which are all equally likely and have the same number of cases
      are):
      \begin{itemize}
        \item $A$ appears before $C$ which appears before $E$ (acceptable).

        \item $A$ appears before $C$ but $E$ appears before $C$ (not acceptable).

        \item $C$ appears before $A$ which appears before $E$ (not acceptable).

        \item $C$ appears before $E$ which appears before $A$ (not acceptable).

        \item $E$ appears before $A$ which appears before $C$ (not acceptable).

        \item $E$ appears before $C$ which appears before $A$ (not acceptable).
      \end{itemize}

      Each of these six options have just as many string which obey their rules.
      And all strings obey one of the rules above. Meaning that only a sixth of
      the total number of strings is acceptable. So there are
      \[
        \frac{5!}{6} = 20
      \]
      strings which have $A$ before $C$ which is before $E$.
    \end{solution}
  \end{parts}

  \question[10] Prove that a number is divisible by 3 if and only if  the sum of
  its digits is divisible by 3. (E.g. 27 is divisible by 3 since $2+7 = 9$ is
  divisible by 3).
  \begin{solution}
    ($\Rightarrow$) If a number, $n$, is divisible by 3 then $n = 3k$ for some
    integer $k$. We can write $n$ as multiples of powers of 10:
    \[
      n = a_k 10^k + a_{k-1} 10^{k-1} + \cdots + a_1 10 + a_0
    \]
    This is simply the base 10 expansion of $n$.

    Thus we have the equation:
    \begin{align*}
      a_k 10^k + a_{k-1} 10^{k-1} + \cdots + a_1 10 + a_0 &= 3k\\
      \intertext{Now take both sides modulo 3, noting that $10\mod 3 \equiv 1$}
      a_k + a_{k-1} + \cdots + a_1 + a_0 \mod 3 \equiv 0
    \end{align*}
    Thus the sum of the digits is also a multiple of 3.

    ($\Leftarrow$) If the sum of the digits of a number is a multiple of three
    we can do the above proof backwards, given that the sum is a multiple of 3
    multiplying each term by 1 does nothing to affect the sum, 10 is equivalent
    to 1 modulo 3 thus $n$ is also a multiple of 3.
  \end{solution}

  \question[15] All numbers have some divisibility rule associated with them.
  There are a couple for 7. The one we focus on here is that a number $n$ is
  divisible by 7 if and only if when you subtract 2 times the least significant
  digit from the number without its least significant digit the result is also
  divisible by 7. E.g. 14 is divisible by 7 since $1 - 4\cdot 2 = -7$ is
  divisible by 7. 343 is divisible by 7 since $34 - 3\cdot 2 = 28$ which is
  divisible by 7. Prove this.
  \begin{solution}
    To prove this we do a similar trick as the previous problem: If a number is
    divisible by 7 we can write $n = 7k$ for some integer $k$. Write $n$ in a
    particular form:
    \[
      n = a_1 10 + a_0
    \]
    E.g. we could write $30248 = 3024\cdot 10 + 8$

    Then our equation becomes:
    \begin{align*}
      a_1 10 + a_0 &= 7k
      \intertext{Now take both sides modulo 7, noting that $10\mod 7 \equiv 3$}
      a_1 3 + a_0 \mod 7 \equiv 0
    \end{align*}
    This gives a rule that we could split a number up by taking off the last
    digit, multiplying the number made up of the large digits by 3 then add the
    remaining digit, e.g. 343 is divisible by 7 since $34\cdot 3 +3 = 105$ and
    105 is divisible by 7 since $10\cdot 3 + 5 = 35$ which is divisible by 7. 

    This rule however doesn't scale well at all. Imagine needing to check
    whether 13983, were divisible by 7 in your head, multiplying 1398 by 3 will
    be a lot of work.

    Thus we keep working with our equation above. It would be convenient if we
    could get rid of the multiple of the higher digits. Thus we need to multiply
    by the inverse of 3 modulo 7. $3^{-1} \mod 7 \equiv 5$. Thus our rule
    becomes:
    \[
      a_1 + a_0 5 \mod 7 \equiv 0
    \]
    Which is better, we're only multiplying a small number by 5. However we can
    actually make this easier by remembering that $5\equiv -2\mod 7$ thus our
    equation is $a_1 -2 a_0 = 7k$.

    Thus our algorithm is to split the least significant digit off the number,
    multiply that by 2 and subtract it from the number without the least
    significant digit.
  \end{solution}

  \question For the following questions find the number of (unordered) five-card
  poker hands, selecting from an ordinary 52-card deck, having the properties
  indicated.
  \begin{parts}
    \part[5] Containing four of a kind, that is, four cards of the same
    denomination.
    \begin{solution}
      In order to get a four of a kind we must first choose which of the 13
      denominations we will duplicate: there are $\binom{13}{1}$ options for
      this.

      Then we must choose which of the remaining 12 denominations we will allow
      for our remaining card: $\binom{12}{1}$ ways to do this.

      Then we must choose which suit it will be, there are $\binom{4}{1}$ ways
      this could fall.

      Thus there are $\binom{13}{1} \cdot \binom{12}{1} \cdot \binom{4}{1} =
      624$
    \end{solution}

    \part[5] Containing cards of exactly two suits
    \begin{solution}

      This one is complicated we need to first over count and then eliminate
      some possibilities.

      Initially we count there are $\binom{4}{2}$ ways to count which suits are
      represented and then there are $\binom{26}{5}$ ways of choosing our five
      cards from among the 26 possible (only 2 suits allowed remember!). This
      seems good, BUT this count also includes the possibility that we selected
      all the same suit. Thus we need to subtract some values.

      Thus we need to count how many hands have cards from one suit, essentially
      a flush. There are $\binom{4}{1}$ ways of choosing which suit and
      $\binom{13}{5}$ ways of choosing those cards.

      Actually when we counted the possible hands with two suits we actually
      counted each of the ways with one suit 3 times (can you see why?)

      Thus there are:
      \[
        \binom{4}{2}\cdot\binom{26}{5} - 3\cdot\binom{4}{1}\cdot\binom{13}{5} =
        379,236
      \]
      ways of getting exactly 2 suits.
    \end{solution}

    \part[5] Containing two of one denomination, two of another denomination and one of a third denomination.
    \begin{solution}
      In order to choose the first two we first see how many ways we have to
      choose a denomination: $\binom{13}{1}$.

      Then we must select two of this denomination: $\binom{4}{2}$.

      Then we must select a denomination for the second two: $\binom{12}{1}$.

      And which cards are selected: $\binom{4}{2}$.

      Finally we select the third denomination: $\binom{11}{1}$.

      And the card of that denomination: $\binom{4}{1}$.

      Altogether we have:
      \[
        \binom{13}{1}\cdot \binom{4}{2}\cdot \binom{12}{1} \cdot
        \binom{4}{2}\cdot \binom{11}{1}\cdot \binom{4}{1} = 247,104
      \]
      ways of forming hands of that type.
    \end{solution}
  \end{parts}

  \question The following questions deal with selecting a committee from a club
  consisting of six distinct men and seven distinct women.
  \begin{parts}
    \part[4] In how many ways can we select a committee of three men and four
    women?
    \begin{solution}
      $\binom{6}{3}\cdot \binom{7}{4} = 700$.
    \end{solution}

    \part[4] In how many ways can we select a committee of four persons that has
    at least one man?
    \begin{solution}
      $\binom{13}{4} = 715$ ways to choose a committee of four people, but we
      need to eliminate all of the ones with no men which is $\binom{7}{4} =
      35$ so there are 680 ways.
    \end{solution}

    \part[6] In how many ways can we select a committee of four persons that has
    at most one man?
    \begin{solution}

      From the previous problem we have seen there are 35 ways to select a
      committee with no man. Now just count the ways to choose a committee with
      one man:

      $\binom{6}{1}\cdot \binom{7}{3} = 210$ ways. Thus there are 245 ways to
      select a committee with at most one man.
    \end{solution}
  \end{parts}
\end{questions}
\end{document}
