%Thedore Ian Martiny
%Summer 2017 Homework 3

\documentclass[addpoints]{exam}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\rationals}{\mathbb{Q}}
\renewcommand{\baselinestretch}{1.5}
%\setlength{\textwidth}{16cm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wasysym}
\usepackage{xcolor}
 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\definecolor{Gray}{gray}{0.85}
 
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\newcolumntype{a}{>{\columncolor{Gray}}c} 
\lstset{style=mystyle}
\begin{document}
\singlespacing

\begin{center}
  {\large\textbf{CSCI 2824 - Discrete Structures}}

  {\large\textbf{Homework 3}}
\end{center}

You MUST show your work. If you only present answers you will receive minimal 
credit. This homework is worth \numpoints pts.

\textbf{Due: Wednesday June 28}

\begin{questions}
  \question[4] For each of the following determine the number of elements in the
  given set. 

  \begin{parts}
    \part $\{\}$
    \begin{solution}
      0
    \end{solution}

    \part $\{\{\},\{\{\}\}\}$.
    \begin{solution}
      2
    \end{solution}

    \part $\{a,b,\{\},\{\{\{\}\}\}\}$
    \begin{solution}
      4
    \end{solution}

    \part $\{a,b,\{a,b\},\{a,c\},\{a\}\}$
    \begin{solution}
      5
    \end{solution}
  \end{parts}

  \question[5] For the following pairs of sets, determine which operator goes 
  between the sets to make a true statement: $\in$, $\ni$, $\subseteq$, 
  $\supseteq$, or none.

  \begin{parts}
    \part $\{1,2\}$, $\{1,2,\{1,2\}\}$
    \begin{solution}
      $\{1,2\}\in\{1,2,\{1,2\}\}$. However $\{1,2\}\subseteq\{1,2,\{1,2\}\}$ as
      well.
    \end{solution}

    \part $\{1,2\}$, $\naturals$
    \begin{solution}
      $\{1,2\}\subseteq\naturals$
    \end{solution}

    \part $\{\naturals,\reals\}$, $\{\reals\}$
    \begin{solution}
      $\{\naturals,\reals\} \supseteq \{\reals\}$
    \end{solution}

    \part $\{\reals\}$, $\{1,3,4\}$
    \begin{solution}
      There is no relation between these two sets.
    \end{solution}

    \part $\reals$, $\{1,\pi, \sqrt{2}, \sqrt{-1}\}$
    \begin{solution}
      There is no relation between these two sets ($\sqrt{-1}$ is not a real 
      number).
    \end{solution}
  \end{parts}

  \question[5] For each of the following determine whether or not it is a 
  function, if not explain why not.

  \begin{parts}
    \part $f:A\to B$ where $A = \{1,2,3,4,5\}$ and $B = \{b,x,t,m,z,y,a\}$ given
    by the following set $\{(1,a),(4,b),(2,b)(5,t),(2,a)\}$.
    \begin{solution}
      No, $2\mapsto b$ and $2\mapsto a$.
    \end{solution}

    \part $g:\reals\to\reals$ given by $g(x) = \tan(x)$.
    \begin{solution}
      No. $\tan\left( \frac{\pi}{2} \right)$ has no value in $\reals$.
    \end{solution}

    \part $h: \naturals \to \integers^{>0}$ given by $h(x) = x-1$
    \begin{solution}
      No, 0 and 1 do not map to anything.
    \end{solution}

    \part $k: A\to B$ where $A = \{18,38,485,382385,25\}$ and 
    $B = \{1,2,3,4,5\}$ given by the following set $\{(18,1),(38,1),(285,1),
    (382385,1),(25,1)\}$.
    \begin{solution}
      No. $(285,1)$ makes no sense since $285\not\in A$. Additionally $485\in A$
      maps to nothing.
    \end{solution}

    \part $l: \reals\to\reals$ given by $l(x) = \log(|x|)$.
    \begin{solution}
      No, $l(0)$ is not defined in the reals.
    \end{solution}
  \end{parts}

  \question[5] Prove that if $X\subseteq Y$ then $X\cap Z \subseteq Y\cap Z$ for
  all sets $X,Y,Z$. 
  \begin{solution}
    Let $X,Y,Z$ be sets such that $X\subseteq Y$. To show $X\cap Z \subseteq
    Y\cap Z$ we take an arbitrary element in the first set and show it is also
    in the second set. So let $r \in X\cap Z$, this tells us that $r\in X$
    \textbf{AND} $r\in Z$. Since $X\subseteq Y$ we have that $r\in Y$. So $r\in
    Y$ and $r\in Z$ so that $r\in Y\cap Z$. Thus every element of $X\cap Z$ is
    in $Y\cap Z$ so $X\cap Z \subseteq Y\cap Z$.

    \qed
  \end{solution}

  \question[5] Prove that $\mathcal{P}(X) \cup \mathcal{P}(Y) \subseteq 
  \mathcal{P}(X\cup Y)$. Are they equal? If not give a counterexample.
  \begin{solution}
    Again we choose an arbitrary element of the first set and show it is in the
    second: Let $r\in \mathcal{P}(X)\cup \mathcal{P}(Y)$. This means that $r$ is
    a \emph{set}. In particular $r\subseteq X$ or $r\subseteq Y$. In either case
    $r\subseteq X\cup Y$. Which means that $r\in\mathcal{P}(X\cup Y)$.

    \qed

    These sets are NOT always equal, for example choose $X = \{1,2\}$ and $Y =
    \{a,b,c\}$ then $X\cup Y = \{1,a,2,b,c\}$. So $\{1,a\} \in \mathcal{P}(X\cup
    Y)$ but $\{1,a\}\not\in \mathcal{P}(X)\cup \mathcal{P}(Y)$ since $a\not\in
    X$ and $1\not\in Y$.
  \end{solution}
  
  \question[20] For the following statements either give a proof or a 
  counterexample. The sets $X,Y,Z$ are subsets of a universal set $U$. Counter 
  examples must also include the definition for $U$.
  \begin{parts}
    \part For all sets $X$ and $Y$, either $X\subseteq Y$ or $Y\subseteq X$.
    \begin{solution}
      This is false. Let $U= \{1,2,3\}$ and $X = \{1\}$ and $Y = \{2\}$ then $X\not\subseteq Y$ and $Y \not\subseteq X$. 
    \end{solution}

    \part $\overline{Y\backslash X} = X \cup \overline{Y}$
    \begin{solution}
      This is true. Let $U$ be any universal set and $X,Y\subseteq U$. We show
      the two given sets are subsets of each other. For reference we expand $
      \overline{Y\backslash X}$, $a\in \overline{Y\backslash X}$ means that
      $a\in U$ but $a\not\in Y\backslash X$. However $a$ may be in $X$.
      
      ($\subseteq$) If $a \in \overline{Y\backslash X}$ then that is saying that
      $a\in U \backslash (Y\backslash X)$. That is $a$ is in $U$ and $a$ is not 
      in $Y\backslash X$. If $a\in X$ then we are good, if it is not then that 
      means that $a\in U$, but also we know for a fact that $a\not\in Y$, so
      that $a\in U\backslash Y$ or $a\in \overline{Y}$.

      ($\supseteq$) If $a \in X \cup \overline{Y}$ we have two cases, either 
      $a\in X$ in which case $a\in \overline{Y\backslash X}$ by above. If $a \in
      \overline{Y}$ this means that $a\in U\backslash Y$, or $a\in U$ and not in
      $Y$. So that $a\in \overline{Y\backslash X}$, a larger set than 
      $\overline{Y}$.

      \qed
    \end{solution}

    \part $X \cup \left(Y\backslash Z\right) = \left(X \cup Y\right)\backslash 
    \left(X\cup Z\right)$
    \begin{solution}
      This is not true. Consider $U = \{1,2,3,4,5\}$, $X= \{1,2,3\}$, $Y=\{4,5\}$,
      $Z= \{3,5\}$. Then $X\cup \left(Y\backslash Z\right) = \{1,2,3,4\}$ but 
      $\left(X \cup Y\right)\backslash \left(X\cup Z\right) = \{4\}$.

      \qed
    \end{solution}

    \part $X \times \left( Y\cup Z\right) = \left(X\times Y\right) \cup \left(
    X\times Z\right)$
    \begin{solution}
      This is true. Let $U$ be a universal set, and $X,Y,Z\subseteq U$. 

      ($\subseteq$) If $a\in X\times \left(Y\cup Z\right)$ then $a$ is of the
      form $(x,b)$ where $x\in X$ and $b\in Y\cup Z$. Thus $a\in X\times Y$ or 
      $a\in X\times Z$. Thus $a\in \left(X\times Y\right) \cup \left(X\times
      Z\right)$

      ($\supseteq$) If $a\in \left(X\times Y\right) \cup \left( X\times
      Z\right)$ then $a$ is of the form $(x,b)$ where $x\in X$ but depending on
      where $a$ is either $b\in Y$ or $b\in Z$. Thus $a\in X\times \left(Y\cup
      Z\right)$.

      \qed
    \end{solution}
  \end{parts}

  \question[3] For the following problems a function definition is given. You 
  must describe a domain and co-domain that ensures it is actually a function.
  The co-domain you provide need not be precisely the range.

  \begin{parts}
    \part $m(x) = \log(x)$.
    \begin{solution}
      Define $m: \reals^{>0} \to \reals$
    \end{solution}

    \part $n(x) = 12$
    \begin{solution}
      We can define the domain to be any set and the co-domain to be any set 
      containing 12.
    \end{solution}

    \part $o(x)$ defined by the set $\{(1,2),(\pi,3),(12,3)\}$
    \begin{solution}
      We can use a domain to be any set containing $1,\pi,12$ and the
      co-domain to be any set containing $2,3$.
    \end{solution}
  \end{parts}

  \question[12] For the following functions determine whether they are 
  one-to-one or onto or both or neither. 
  \begin{parts}
    \part $f:\integers \to\integers$, $f(n) = n+1$
    \begin{solution}
      This function is injective, if $m\not= n$ then $m+1\not= n+1$.

      This is surjective, given $m\in \integers$ then $f(m-1) = m$.

      Thus this is a bijection.
    \end{solution}

    \part $g:\integers\to\integers$, $g(n) = \lceil \frac{n}{2}\rceil$.
    \begin{solution}
      This function is not one-to-one, $g(0) = 0 = g(1)$.

      This function is surjective, given $m\in \integers$ then $g(2m) = \lceil \frac{2m}{2} \rceil = \lceil m \rceil = m$.
    \end{solution}

    \part $h: \integers\times\integers \to \integers$, $h(m,n) = m-n$.
    \begin{solution}
      This function is not injective, $h(0,0) = 0 = h(2,2)$.

      This function is surjective, given $m\in integers$ then $h(m,0) = m$.
    \end{solution}

    \part $j: \integers\times\integers \to \integers$, $j(m,n) = m^2 + n^2 + 2$.
    \begin{solution}
      This function is not injective $j(0,1) = 3 = j(1,0)$.

      This function is not surjective there is no $(m,n)\in\integers\times\integers$ to result in a negative integer.
    \end{solution}
  \end{parts}

  \question[5] Give examples of functions from $\naturals$ to $\naturals$ that
  are:
  \begin{parts}
    \part one-to-one, but not surjective
    \begin{solution}
      Choose $f:\naturals\to\naturals$ defined by $f(n) = 2n$. This is not
      surjective because nothing maps to 3. It is injective:

      Choose $n\not=m \in \naturals$ then $2n\not= 2m$.

      \qed
    \end{solution}

    \part surjective but not injective
    \begin{solution}
      Choose $g:\naturals\to\naturals$ defined by:
      \[
        g(n) = \begin{cases}
          0 & \text{ if } n = 0, 1\\
          n-1 & \text{ otherwise}
        \end{cases}
      \]
      Then $g$ is clearly not injective, both $0$ and $1$ map to $0$. But is
      surjective:

      Choose $m\in\naturals$ if $m = 0$ then $f(0) = 0$ otherwise $f(m+1) = m$.
    \end{solution}

    \part injective and surjective (but not the identity function)
    \begin{solution}
      Choose $h:\naturals\to\naturals$ defined by:
      \[
        h(n) = \begin{cases}
          1 & \text{ if } n = 0\\
          0 & \text{ if } n = 1\\
          n & \text{ otherwise}
        \end{cases}
      \]
      This is not the identity function since $f(1) = 0 \not=1$. But it is
      surjective: choose $m\in\naturals$ if $m= 0$ then $f(1) = m$, if $m = 1$
      then $f(0) = 1$, otherwise $f(m) = m$.

      It is also injective, if $n\not=m$ then we can examine by cases, if
      necessary, to see that $f$ does not map these different values to the same
      value.
    \end{solution}

    \part neither injective nor surjective.
    \begin{solution}
      Choose $j:\naturals\to\naturals$ defined by $j(n) =1$. This is not
      surjective, nothing maps to 2. And this is not injective, everything maps
      to the same value.
    \end{solution}
  \end{parts}

  \question[11] Prove that the function $f:\integers^{>0}\times\integers^{>0}\to
  \integers^{>0}$ defined by $f(m,n) = 2^m\cdot 3^n$ is injective but not 
  surjective.
  \begin{solution}
    This is injective: If $(m,n) \not= (a,b)$ then we have cases:

    \textbf{Case 1:} $m\not= a$ then $2^m \not= 2^a$. Then no matter how many
    multiples of three are multiplied on we have that $2^m \cdot 3^n \not= 2^a
    3^b$.

    \textbf{Case 2:} $n \not= b$ then $3^n \not= 3^b$. Again no matter how many
    multiples of two are multiplied on we have that $2^m \cdot 3^n \not= 2^a
    3^b$.

    Thus in any case we have that $f(m,n) \not= f(a,b)$ so we have an injective
    function.

    However this function is not surjective. Only $m = 0$ can make $2^m = 1$ and
    only $n=0$ can make $3^n = 1$ but 0 is not a valid value for $m$ or $n$.
    Thus this function is not surjective.

    Other examples are that no multiplication of multiples of $2$ and $3$ can
    result in $5$.

    \qed
  \end{solution}

  \question[10] Solve the following recurrence relations (provide a closed form
  solution):
  \begin{parts}
    \part $a_n = -3a_{n-1}$, $a_0 = 4$
    \begin{solution}
      Solving this by reducing terms:
      \begin{align*}
        a_n &= -3 a_{n-1}\\
        &= -3 (-3) a_{n_2} = 9 a_{n-2}\\
        &= -3 (-3) (-3) a_{n-3} = - 27 a_{n-3}\\
        &= \vdots\\
        &= (-3)^{k} a_{n-k}\\
        \intertext{Letting $k = n$:}
        &= (-3)^{n} a_0\\
        &= (-3)^{n} 4
      \end{align*}
    \end{solution}

    \part $a_n = a_{n-1} + 1$, $a_0 = 12$
    \begin{solution}
      Again reducing terms:
      \begin{align*}
        a_n &= a_{n-1} + 1\\
        &= a_{n-2} + 1 + 1\\
        &= a_{n-3} + 1 + 1 + 1\\
        &= \vdots\\
        &= a_{n-k} + k\\
        \intertext{Letting $k = n$:}
        &= a_0 + n\\
        &= 12 + n
      \end{align*}
    \end{solution}
  \end{parts}

  \question[10] Solve the following recurrence relations (provide a closed form
  solution):
  \begin{parts}
    \part $a_n = 6a_{n-1} - 8a_{n-2}$, $a_0 = 1$, $a_1 = 0$.
    \begin{solution}
      This problem leads to the characteristic equation:
      \[
        x^2 - 6x + 8 = 0
      \]
      which factors as:
      \[
        (x-4)(x-2) = 0
      \]
      and so the solutions are $x =2,4$. Thus the general form of our solutions 
      is:
      \[
        a_n = A2^n + B 4^n
      \]
      Using our initial conditions we get the two equations:
      \begin{align*}
        1 &= A + B\\
        0 &= 2A + 4B
      \end{align*}
      we can solve for $A$ using the second equation: $A = -2B$ and plugging 
      into the first equation we get:
      \begin{align*}
        1 &= -2B + B\\
        1 &= -B\\
        B &= -1
      \end{align*}
      Which gives that $A = 2$. Thus our solution is:
      \[
        a_n = 2(2^n) + (-1)4^n
      \]
      or 
      \[
        a_n = 2^{n+1} + (-1) 4^n
      \]
    \end{solution}

    \part $a_n = 2a_{n-1} + 8 a_{n-2}$, $a_0 = 4$, $a_1 = 10$.
    \begin{solution}
      This problem leads to the characteristic equation:
      \[
        x^2 - 2x - 8 = 0
      \]
      which factors as:
      \[
        (x-4)(x+2) = 0
      \]
      and so the solutions are $x = -2,4$. Thus the general form of our solutions is:
      \[
        a_n = A(-2)^n + B 4^n
      \]
      Using our initial conditions we get the two equations:
      \begin{align*}
        4 &= A + B\\
        10 &= -2A + 4B
      \end{align*}
      If we add two times the first equation to the second equation we get:
      \begin{align*}
        18 &= 0A +  6B\\
        &= 6B\\
        B &= 3
      \end{align*}
      Which gives that $A = 1$. Thus our solution is:
      \[
        a_n = (-2)^n + (3)4^n
      \]
    \end{solution}
  \end{parts}

  \question[5] Give an example of of two uncountable sets $A$ and $B$ such that
  $A\cap B$ is:
  \begin{parts}
    \part finite
    \begin{solution}
      Choose $A = [1,2]$ and $B = [2,3]$, since $A$ and $B$ are intervals,
      subsets of $\reals$ then they are uncountable. But $A\cap B = \{2\}$ which
      is finite.
    \end{solution}

    \part countably infinite
    \begin{solution}
      Choose $A = \reals \times \rationals$ and $B = \rationals \times \reals$
      then both $A$ and $B$ are uncountable since they contain $\reals$, which
      is uncountable, but $A\cap B = \rationals \times \rationals$ which is
      countable infinite.
    \end{solution}

    \part uncountably infinite
    \begin{solution}
      Choose $A = [1,3]$ and $B = [2,4]$ then $A$ and $B$ are uncountable, since
      they are interval subsets of $\reals$. $A\cap B = [2,3]$ which is also an
      interval subset of $\reals$ and is thus uncountable.
    \end{solution}
  \end{parts}
\end{questions}
\end{document}
